RadLib: a radiative heat transfer model library for CFD

### 2020-11-17 manuscript/manuscript.tex
Cleaned up motivation section

% Brain dump for first draft
Why did we make this? Radiation is hard to deal with in CFD. It's complicated. You can't really get analytic solutions to practical problems. In lots of cases, that's fine because radiation can be safely neglected; it's not usually the dominant mode of heat transfer. However, we do combustion CFD. Radiation isn't always important to combustion problems, but when it is, simulations can get grossly inaccurate. So we need a nice way to account for radiation in our CFD simulations.

When is radiation important to combustion simulations? In high-temperature systems, radiation can often be neglected in the beginning and middle of the simulation (assuming, say, a jet flame or a counterflow flame or something like that). At those times (which I'm going to call "early flame"), convective heat transfer dominates pretty heavily. However, radiation becomes more important for late-stage flame phenomena, either after some time has gone by (radiation time scales?), or physically high up in a flame where convection doesn't dominate as much as below, or just in areas that are relatively far away from the flame sheet itself where the reactions are happening. There's also soot, which involves radiation heavily, but soot particles are so big and slow that their time and length scales (and their radiative time scales) mean that radiation doesn't become important until late in the flame evolution.

So what's the problem? Why hasn't this been done already? Radiation is complicated. So is combustion. Combustion in particular can be so complex that direct simulations are too computationally expensive for us to simulate configurations that are practical for engineering systems. Direct simulations are usually used as a research tool. Other modeling approaches exist (LES, ODT, etc.) that lower the computational cost and allow us to simulate practical things. So far so good.

Radiation is a little like combustion in that its core mechanisms are complex, physically and mathematically. In addition, it's directional AND depends on the wavelength of the energy involved (other heat transfer doesn't have the wavelength dependence); its governing equations integrate over direction and wavelength, which makes things extra complicated. RADIATION GOVERNING EQUATIONS HERE.  Convective heat transfer follows easy rules, but radiation doesn't. There are simple systems in which we can boil things down to analytic solutions, but most of the time, the simple equations don't apply, and that's often true in combustion systems where there are so many different length and time scales involved. The fundamental equations of radiative heat transfer are big and mathematically complex; they don't have analytic solutions except in the simplest geometries. So what do we do? One of two things. One, we do ray tracing (sometimes referred to as Monte Carlo simulations. These can be extremely accurate, but super computationally expensive. Essentially the equivalent of a direct numerical solution. Two, we simplify the the equations. We make assumptions. And so on. This is potentially less accurate, but more practical. By simplifying the fundamental governing equations, we create models that apply to various situations, systems, and geometries (i.e. black body assumptions, directional assumptions, etc.). There are models developed for CFD and specific combustion systems. These are the models that we've put into practice here in such a way that they're easy to apply to various systems. Modular organization for this purpose.

So far there isn't an easy way to access and use radiation models. There are simple ones in Cantera (check this, there might not be any), but they make too many assumptions or don't work well for combustion or what have you. RadLib is a radiation library of models that you can apply to any simulation type, and we'd like to add it to Cantera, too. Basically, there isn't a reliable way to do radiation calculations in CFD simulations without coding the models yourself, which is extra hard because they're complicated and hard. So we've done it for you and put them in a library that easy for anyone to use.

%Author guidelines for this section:
%\begin{itemize}
% \item Introduce the scientific background and the motivation for developing the software. 
% \item Explain why the software is important, and describe the exact (scientific) problem(s) it solves.
% \item Indicate in what way the software has contributed (or how it will contribute in the future) to the process of scientific discovery; if available, this is to be supported by citing a research paper using the software.
% \item Provide a description of the experimental setting (how does the user use the software?).
% \item Introduce related work in literature (cite or list algorithms used, other software etc.).
%\end{itemize}

%Author guidelines for this section
%\begin{itemize}
% \item Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.
% \item  Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.
% \item Present the major functionalities of the software.
%\end{itemize}

@@

When we refer to the Plank Mean model, what we're using is actually the Planck Mean (PM) absorption coefficients, calculated from the correlations given on the TNF workshop site \citep{Smith_2003}. Their correlations (temperature dependent) are based on the RADCAL model in \citep{Grosshandler_1993}. The TNF radiation model is also documented in \citep{Barlow_2001}.

This model is commonly used because it's relatively simple, easy to implement, low in computational expense, and provides reasonably accurate results in many cases.

Quote from TNF site page: "The characteristics of some flames selected for the workshop are such that a model based upon the assumption of optically thin radiative heat loss should yield reasonable accuracy. This has been demonstrated for the simple hydrogen jet flames \cite{Barlow_1999}. However, there is some evidence that the optically thin model significantly over predicts radiative losses from the CH4 flames in the TNF library, due to strong absorption by the 4.3-micron band of CO2 \cite{Frank_2000,Zhu_2002,Coelho_2002}." Basically, it works well for flames that are actually optically thin, but not so well for flames that aren't. This is especially true of sooting flames.

@@

RadLib is an object-oriented C++ class library, including both C++ and Python interfaces. The RadLib package contains five subdirectories (upon initial download): \texttt{source} contains the RadLib source code; \texttt{build} contains installation files; \texttt{examples} contains instructive example cases, including a simple interface and solver for a parallel planes geometry; \texttt{data} contains ALBDF data tables required for the RCSLW model; and \texttt{docs} contains files used to generate code documentation with Doxygen (optional).

@@

Several examples are presented to illustrate the behavior of the models. The examples show heat flux $q$ or volumetric heat source $Q$ one one-dimensional configurations with varying gas compositions and temperatures. We compare the PM, WSGG, and RCSLW models for each example. The examples are taken from those presented by Solvojov et al., (2017) \cite{Solovjov_2017}, and Bordbar et al., 2020 \cite{Bordbar_2020}; the number corresponds to the example in the respective referece. A ray-tracing code is used to solve the radiative transport equation between two parallel plates. 
Table~\ref{t:examples} summarizes the cases. Example S1 is a hot slab next to a cold slab where the cold thickness varies; Example S2 is similar but isothermal with a \emph{thick} slab of high CO$_2$ next to a \emph{thin} slab of low CO$_2$ of varying thickness; Example S3 has parabolic temperature and H$_2$O profiles; Example S4 has a triangular temperature profile between equally-spaced isothermal regions; Example S5 is a half-sinusoid decreasing in temperature from 1500 to 500 K; and Example B3 has symmetric temperature and H$_2$O profiles with central peaks of 1800 K and 1, respectively (with $y_{CO2}=1-y_{H2O}$). In each case, comparison is made to the line-by-line (LBL) data presented in the references. 
%
\begin{table}
    \caption{Summary of example cases presented. S1-S5 are from \cite{Solovjov_2017}; B3 is from \cite{Bordbar_2017}. All cases have $P=1$ atm, black walls.}
    \label{t:examples}
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c l l c c c}
        \hline
        Example & T(K)                           & $y_{H2O}$  $y_{CO2}$ (mole frac.)                     & L (m)   & $T_{walls}$ (K)\\
        \hline
        S1 & $T(x<0.5)=2000$; $T(x>0.5)=300$     & $y_{CO2}=0.1$, $y_{H2O}=0.2$                          & 0.5-2.5 & cold, cold     \\
        S2 & T=1000                              & $y_{CO2}(x<0.5)=0.4$, $y_{CO2}(x>0.5)=0.1$            & 0.5-2.5 & cold, cold     \\
           &                                     & $y_{H2O}=0.0$                                         &         &                \\
        S3 & $T(x) = 4000x(L-x)/L^2 + 800$       & $y_{H2O}(x) = 0.8x(L-x)/L^2 + 0.12$                   & 1       & 800, 800       \\
           &                                     & $y_{CO2}=0$                                           &         &                \\
        S4 & middle third triangular to 2500     & $y_{H2O}=0.1$, $y_{CO2}=0$                            & 0.3     & 500, 500       \\
        S5 & $T(x) = 1000 + 500\cos(\pi x/L)$    & $y_{H2O}=0.1$, $y_{CO2}=0$                            & 2       & 1500, 500      \\
        B3 & $T(x) = 400 + 1400\sin(\pi x/L)^2$  & $y_{H2O}(x) = 0.0001 + 0.9999\sin(\pi x/L)^2$         & 1       & 400, 400       \\
           &                                     & $y_{CO2}=1-y_{H2O}$                                   &         &                \\
        \hline
    \end{tabular}
    }
\end{table}
%

>>>>

Modeling radiative heat transfer, and particular radiation absorption coefficients, for CFD is complex and challenging. More specifically, combustion simulations are often complicated by the need for accurate radiation modeling, which can be difficult to implement and computationally expensive relative to simulation cost. As a result, researchers typically manufacture configurations in which radiation can be safely ignored, which can lead to results inapplicable to practical systems, or use oversimplified models and assumptions, which can produce inaccurate results. Unfortunately, these approaches are neither rigorous nor applicable to real engineering systems.

In some cases, other modes of heat transfer dominate the system, and radiation can be safely neglected, but many practical engineering systems do not allow this. Simple jet flames, for example, demonstrate both possibilities: early in a flame's development, convection typically dominates heat transfer and dictates flame behavior; late-stage flame phenomena like soot behavior and flame sheet breakthrough, however, increasingly depend on the magnitude of radiative heat losses and cannot be simulated accurately without a radiation model.

Like combustion, the core mechanisms of radiation are physically and mathematically complex. Unlike other modes of heat transfer, it is governed by a set of integro-differential equations that depend on both direction and wavelength. Only the simplest systems have analytic solutions, and most practical systems require a numerical solution. Ray tracing (Monte Carlo) solutions are very accurate but extremely computationally expensive, and often require high-performance computing resources to execute. Modeling assumptions are common and can reduce simulation cost at the expense of accuracy, depending on their validity for the system in question.

Currently, there is no convenient or easy-to-use access point for CFD radiation models. As such, we present RadLib, a modular library of radiation models that can be applied alongside various simulation tools. RadLib focuses on calculation of radiation absorption coefficients, often the most difficult and time-consuming portion of radiation calculations, which permits its use with various solvers. The models are fully implemented and validated and all use a common interface, allowing researchers easy and convenient access to radiation modeling tools regardless of the field of application.

@@

RadLib's simplest model uses Planck Mean absorption coefficients calculated from the correlations given on the TNF Workshop site \citep{Smith_2003}. The temperature-dependent correlations are based on the RADCAL model in \citep{Grosshandler_1993}. The TNF radiation model is also documented in \citep{Barlow_2001}.

Planck Mean absorption coefficients are commonly used to model radiation, especially in combustion systems, because the model is relatively easy to implement, computationally inexpensive, and reasonably accurate in many cases. The TNF Workshop correlations were developed for use with an optically thin radiation model, which assumes that radiation passes through a medium mostly undisturbed. In cases where the optically thin assumption applies, such as simple hydrogen jet flames, this model and the associated Planck Mean absorption coefficients can produce accurate results \citep{Barlow_1999}. In cases where the optically thin assumption is not reasonable, including many other combustion scenarios, this model does not produce accurate results. For example, it significantly overpredicts radiative losses from the TNF library's CH$_4$ flames \cite{Frank_2000,Zhu_2002,Coelho_2002}.

Because RadLib does not specify case geometry or medium, there is no optically thin assumption inherent in its use of the Planck Mean absorption coefficients. However, the correlations were developed with the intention of use with an optically thin assumption, and, as a result, depend only on local temperature values. RadLib users must take care to consider these limitations before using this model.

@@

RadLib is an object-oriented C++ class library that includes both C++ and Python interfaces. The RadLib package contains five subdirectories (upon initial download): \texttt{source} contains the RadLib source code; \texttt{build} contains installation files; \texttt{examples} contains instructive example cases, including a simple interface and solver for a parallel planes geometry; \texttt{data} contains ALBDF data tables required for the RCSLW model; and \texttt{docs} contains files used to generate code documentation with Doxygen (optional).

@@

Several examples are presented to illustrate the behavior of the models. The examples show heat flux $q$ or volumetric heat source $Q$ one one-dimensional configurations with varying gas compositions and temperatures. We compare the PM, WSGG, and RCSLW models for each example. The examples are taken from those presented by Solvojov et al., (2017) \cite{Solovjov_2017}, and Bordbar et al., 2020 \cite{Bordbar_2020}; the number corresponds to the example in the respective referece. A ray-tracing code is used to solve the radiative transport equation between two parallel plates. 
Table~\ref{t:examples} summarizes the cases. Example S1 is a hot slab next to a cold slab where the cold thickness varies; Example S2 is similar but isothermal with a \emph{thick} slab of high CO$_2$ next to a \emph{thin} slab of low CO$_2$ of varying thickness; Example S3 has parabolic temperature and H$_2$O profiles; Example S4 has a triangular temperature profile between equally-spaced isothermal regions; Example S5 is a half-sinusoid decreasing in temperature from 1500 to 500 K; and Example B3 has symmetric temperature and H$_2$O profiles with central peaks of 1800 K and 1, respectively (with $y_{CO2}=1-y_{H2O}$). In each case, comparison is made to the line-by-line (LBL) data presented in the references. 
%
\begin{table}
    \caption{Summary of example cases presented. S1-S5 are from \cite{Solovjov_2017}; B3 is from \cite{Bordbar_2020}. All cases have $P=1$ atm, black walls.}
    \label{t:examples}
    \centering
    \resizebox{\textwidth}{!}{
    \begin{tabular}{c l l c c c}
        \hline
        Example & T(K)                           & $y_{H2O}$  $y_{CO2}$ (mole frac.)                     & L (m)   & $T_{walls}$ (K)\\
        \hline
        S1 & $T(x<0.5)=2000$; $T(x>0.5)=300$     & $y_{CO2}=0.1$, $y_{H2O}=0.2$                          & 0.5-2.5 & cold, cold     \\
        S2 & T=1000                              & $y_{CO2}(x<0.5)=0.4$, $y_{CO2}(x>0.5)=0.1$            & 0.5-2.5 & cold, cold     \\
           &                                     & $y_{H2O}=0.0$                                         &         &                \\
        S3 & $T(x) = 4000x(L-x)/L^2 + 800$       & $y_{H2O}(x) = 0.8x(L-x)/L^2 + 0.12$                   & 1       & 800, 800       \\
           &                                     & $y_{CO2}=0$                                           &         &                \\
        S4 & middle third triangular to 2500     & $y_{H2O}=0.1$, $y_{CO2}=0$                            & 0.3     & 500, 500       \\
        S5 & $T(x) = 1000 + 500\cos(\pi x/L)$    & $y_{H2O}=0.1$, $y_{CO2}=0$                            & 2       & 1500, 500      \\
        B3 & $T(x) = 400 + 1400\sin(\pi x/L)^2$  & $y_{H2O}(x) = 0.0001 + 0.9999\sin(\pi x/L)^2$         & 1       & 400, 400       \\
           &                                     & $y_{CO2}=1-y_{H2O}$                                   &         &                \\
        \hline
    \end{tabular}
    }
\end{table}
%
